### 2016-06-16 writeups/psycMethodsResponse2/response.tex

\bigskip
\bigskip

\begin{questions}

\question Primary concern: Papers on predictive modeling typically avoid confidence intervals and instead seek to optimize out-of-sample RMSE or R2 based on training versus test samples (e.g., k-fold cross-validation) or sometimes by resampling (bootstrapping); see overviews of predictive modeling such as http://www-bcf.usc.edu/~gareth/ISL/ and http://appliedpredictivemodeling.com/.  The general concern here is that people reading your paper alongside other 'big data' papers may view your confidence interval approach unusual or outdated; the more specific concern here is overfitting the model to the data (making confidence intervals too confident).

\begin{solution}
Added a paragraph in the overall method section to address this concern.
The text is included below:

``This research centers around evaluating the differences in model accuracy when various components are added, removed, or altered. When evaluating the models, overall model accuracy for a specific model and specific dataset is less important than how that accuracy deviates from a comparative model on the same dataset. Therefore, model overfitting is less of a concern in this case as long as all candidate models have equal opportunity to be slightly overfit. Nonetheless, we did ensure that the models would not be substantially overfit in two ways: The posts used to build the co-occurrence matrices were never included as posts to calibrate weights of model terms. This is particularly important since the context term can easily latch on to a post with exactly the same co-occurrence structure as a post that was used to build the underlying matrix. Second, for each dataset sample, the logistic regression model that calibrated the model terms was fed with a large number of posts (500). Each model only had a few terms to calibrate (e.g., prior, context component), and there were on average 1,300 tags per sample that could be used by the logistic regression model for calibration, so the risk of overfitting was low. Further, Stanley and Byrne (2013) showed that the out-of-sample model and within-sample model accuracy difference for a similar model on a similarly sized dataset for StackOverflow is small (2\% accuracy difference). We were therefore more concerned with the risk of model overfitting by using fewer and more similar dataset samples than from optimizing the weights of a few model terms when those weights were calibrated from thousands of data points. This is why we focused on evaluating the models on multiple dataset samples (particularly the 60 different dataset samples used for Twitter) rather than to perform a k-fold cross validation or out-of-sample validation technique for the weights from the logistic regression.''

@@

\end{solution}

\question Lesser concern: Per Rev 3 para 3, consider how prediction might increase when methods dynamically afford an increase in model dimensionality whenever the size of a corpus increases (this could be considered in terms of differences in sizes across corpora and/or across users). It is fine if you simply point this out as future research in your concluding statements, although an example could be incorporated if you so desired. Your choice.

\begin{solution}
This work was already done but was not yet included in the submission.
I have added it to the paper and the text is included below for convenience.

``Another parameter in the random permutation model is the number of rows used to represent all of the different words. Each word’s environment vector is represented by two randomly placed ones and two negative ones across the rows. All that is required is that the four positions and signs of each of these environment vectors is unique. But this means multiple environment vectors can overlap when looking at a single row. This is a lossy form of compression, and allows the number of rows to represent the words to be much less than the total number of words. As the number of rows decreases, the amount of crosstalk between the environment vectors increases, compression increases, and the amount of information stored decreases. As a third architectural concern, the effect of compression for the random permutation model was examined for Twitter and StackOverflow.

Method. Three different row-size values were tested for the random permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix. A low value (100), a more standard value (2048), and a high value (10,000) of rows were tested. The 2048-row matrix was the default size used for the random permutation model for all other results. A separate co-occurrence representation was built for each of these row values. Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset. Accuracy measurements were the same as were used for previous runs of these datasets: mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.

Results. Results for using different levels of compression for the random permutation model for the StackOverflow dataset are depicted in Figure 13. The plot is also broken out for different size datasets used to build the co-occurrence matrix.

Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy, which shows that compression can influence model accuracy. However, 2048 rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (10,000) has little to no positive impact on accuracy, even for the highest corpus sizes.

Results are similar for Twitter. Figure 14 shows the same negative impact when reducing from 2048 to 100, and little to no impact when increasing from 2048 to 10,000. These results are consistent across all corpus sizes used to build the co-occurrence matrix.

Discussion. The fact that the random permutation model only needs around 2048 rows to represent all of the words in the Twitter and StackOverflow datasets is impressive. Twitter has around 2,383,000 unique words for the 3,000,000 tweets in each of the popular-users datasets. StackOverflow has 5,848,000 unique words for the 1,000,000 posts used to generate the co-occurrence matrix. This means that only 2048 rows are needed to represent most of the variance from all of the words in these co-occurrence matrices. This is likely because word use frequency for StackOverflow and Twitter approximately follows Zipf’s law. Stanley and Byrne (2013) showed that frequency of tag use for StackOverflow approximately follows a Zipf’s law distribution. This means that if the words are rank sorted by number of times they are used, the frequency of word use drops off sharply. Since these low-use words have little to no ability to influence predictions, it makes sense that accuracy does not substantially improve by providing additional dimensions for them.``

@@

\end{solution}

\question Lesser concern: Per Rev 3, para 5 and 7, please incorporate any additional relevant literature (but do not limit yourself to Rev 3 citations; check out what is current and relevant given this quickly changing phenomenon and research area).

\begin{solution}
Added a paragraph in the background section to discuss the types of hashtag-prediction models that were mentioned by Rev 3:

``Other successful models have shown that higher-level processes such as a user’s goals for picking a hashtag, demographics, and specifics about how a hashtag has been used in the past can also influence hashtag use. Lin, Margolin, Keegan, Baronchelli, and Lazer (2013) showed that there are at least two distinct use cases for hashtags on Twitter: to mark the content (traditional tagging) and to contribute to a specific community that has created a hashtag. Yang, Sun, Zhang, and Mei (2012) found that demographics such as gender, age, and location can influence a user’s selected hashtag. Denton, Weston, Paluri, Bourdev, and Fergus (2015) found that the growth and lifetime of a hashtag can be influenced by factors such as how often the hashtag was retweeted and the number of followers for a user that used a particular hashtag. These types of goal-related features, demographic information, and higher-level attributes of previous hashtag use were not directly included in the declarative memory models that were tested for this research. They were not included because it would be difficult to fit these features into a modeling framework that is constrained to use only declarative memory processes, and this research focused on measuring differences in model accuracy for two declarative memory models. Nonetheless, many of these higher-level features may indirectly influence a declarative memory model by influencing the specific tags that a user has chosen in the past, which in turn boosts base-level activation for these previously-chosen tags, which in turn makes it more likely that they are chosen in the future.''

@@

\question That said, declarative memory retrieval is still likely to be a critical part of this process and worth exploration on its own. As such, the paper can still work even without incorporating any of those additional features. Many papers in this area report whether not just whether the tag from the model was predicted, but also whether it was in the top n (10, 25, etc) most likely predictions returned by the model (for one example of this, see ``\#TAGSPACE: Semantic Embeddings from Hashtags'', 2014). One possibility would be to treat the declarative memory portion (modeled here) as generating candidates which other processes consider.

\begin{solution}

There was some text in the beginning of the discussion section that mentioned this possibility.
However it was embedded within parentheses and the other alternatives listed were not as important as this suggestion.
So I broke this specific possibility out of parentheses and mentioned it directly:

``These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process. However, both models are not 100 percent accurate, and the amount of variance left to predict is larger than what would be observed if retrieval noise is included in the model. One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval. For example, it may be the case that the declarative system is used to provide a set of likely tag candidates to higher-level processes to consider for selection.''

### 2016-09-29 writeups/psycMethodsTranslationalAbstract/translationalAbstract.tex
foo

\documentclass[man,floatsintext,donotrepeattitle]{apa6}
%\documentclass[man,donotrepeattitle]{apa6}

% http://tex.stackexchange.com/questions/279/how-do-i-ensure-that-figures-appear-in-the-section-theyre-associated-with
\usepackage{placeins}

\usepackage[nodayofweek]{datetime}% http://ctan.org/pkg/datetime
\usepackage{xfrac}
\usepackage{textcomp}
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}

% Single space after period (Mike prefers)
\frenchspacing

@@

% Changing horizontal space between subsubsection title and text, without loading titlesec package.
% Not loading titlesec because it clashes with apa6 package.
% Took base \@startsection command from apa6 .cls file, and tweaked horizontal spacing. 
% http://tex.stackexchange.com/questions/22576/redefining-sectioning-commands
% http://www.tex.ac.uk/cgi-bin/texfaq2html?label=atsigns
% http://tex.stackexchange.com/questions/60127/subsubsection-description-and-title-not-separated
\makeatletter
\renewcommand{\subsubsection}{%
  \@startsection
  {subsubsection}%
  {3}%
  {\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-.4em}%
  {\normalfont\normalsize\bfseries\addperi}}
\renewcommand{\paragraph}{%
  \@startsection
  {paragraph}%
  {4}%
  {\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-.4em}%
  {\normalfont\normalsize\bfseries\itshape\addperi}}
\makeatother

% ref: http://tex.stackexchange.com/questions/28516/how-to-change-the-title-of-toc
\addto\captionsamerican{% Replace ``english'' with the language you use
  \renewcommand{\contentsname}%
  {Table of Contents}%
}

%\usepackage[compact]{titlesec}
\usepackage{tabu}
\usepackage{amssymb,amsmath}
\usepackage{setspace}

% ref: http://tex.stackexchange.com/questions/48509/insert-list-of-figures-in-the-table-of-contents
\usepackage[nottoc,numbib]{tocbibind}

\usepackage[hyphens]{url}

% ref: http://tex.stackexchange.com/questions/73862/how-can-i-make-a-clickable-table-of-contents
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=black
}
%\hypersetup{linktocpage}

%\usepackage{breakurl}
\usepackage{subfigure}
\usepackage{graphicx}

\usepackage[group-separator={,}]{siunitx}
\newcommand{\numNoZero}[1]{{\sisetup{add-integer-zero=false}\num{#1}}}

\newcommand{\myCI}[4]{{mean difference = \num{#1}, 95\% CI [\num{#2}, \num{#3}]}}

\RequirePackage[l2tabu, orthodox]{nag}
\graphicspath{{./figures/dissertation/}} % Specifies the directory where pictures are stored

\usepackage{csquotes}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber,doi=false,uniquename=false]{biblatex}
\DeclareLanguageMapping{american}{american-apa}

\addbibresource{bibliography.bib}

% Needed to ensure periods are placed at the end of every bibliography entry in the references section
\AtEveryBibitem{\clearfield{doi}}

% http://tex.stackexchange.com/questions/55780/disable-month-in-biblatex-bibliography
\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}

% Make sure only urls that are printed in references section are for webpages
\AtEveryBibitem{
  \ifentrytype{misc}{}{
    \clearfield{url}
  }
}

% ref: http://tex.stackexchange.com/questions/128592/remove-backslashes-from-url-fields-in-bib-entry
\DeclareSourcemap{
  \maps[datatype=bibtex]{
    \map{
      \step[fieldsource=url,
	match=\regexp{\\},
      replace=\regexp{}]
    }
  }
}

% Do not let biblatex change any of the casing in the bibliography file
\DeclareRobustCommand{\MakeSentenceCase*}[1]{{#1}}

% Customer citation for a parencite style but with a few leading words in the parentheses
% Stolen from parencite but with the out parentheses stripped off
% For example: (e.g., Anderson, 1972), (see Salton, 2014)
\DeclareCiteCommand{\noparencite}
  {\usebibmacro{prenote}}
  {\usebibmacro{citeindex}%
   \usebibmacro{cite}}
  {\multicitedelim}
  {\usebibmacro{postnote}}

% If the toc is too detailed, try limiting the depth of printed sections in the TOC:
\setcounter{tocdepth}{2}

\title{Comparing vector-based and Bayesian memory models using large-scale datasets:\protect\\User-generated hashtag and tag prediction on Twitter and StackOverflow}
\shorttitle{Comparing memory models of tag retrieval}
%\shorttitle{}

\author{Clayton Stanley and Michael D. Byrne}
\affiliation{Rice University}

\leftheader{Beitzel}

\keywords{ACT-R declarative memory theory, vector-based models, LSA, machine learning}

@@

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
  By framing the task of choosing a hashtag for a tweet and tagging a post on StackOverflow as a declarative memory retrieval problem,
  two cognitively-plausible declarative memory models were applied to millions of posts and tweets and evaluated on how accurately they predict a user's chosen tags.
  An ACT-R based Bayesian model and a random permutation vector-based model were tested on the large datasets. 
  The results show that past user behavior of tag use is a strong predictor of future behavior.
  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to an entropy-weighting natural language processing method used to attenuate high-frequency words (e.g., articles and prepositions).
  Word order was not found to be a strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

%\note{\today}

\maketitle

%\newpage

\section{Introduction}

With the massive growth of human-created content on social media sites (e.g., Twitter, StackOverflow, Wikipedia) and public access to many of those databases,
it is now possible to evaluate psychological theories of declarative memory on large-scale real-world tasks where the theory can be tested on hundreds of millions to billions of data points.
Without using large datasets it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.

A few researchers have started to use large-scale databases to evaluate how well declarative memory models scale,
for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.

This study expands on previous work that has tested retrieval models on large-scale datasets.
It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutation vector-based) can be improved by testing these models on two real-world tagging tasks.
One primary objective is to compare the random permutation vector-based model to ACT-R's Bayesian declarative memory theory when large datasets are used for each model. 
Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
For example, modifying a vector-based model to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
Large-scale datasets were used throughout the model development and evaluation process.
This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.

As a general methodology for this research, we used a cognitively-constrained exploratory model development process.
Since we worked with large-scale datasets, we had enough data to explore many different modifications simultaneously without the risk of overfitting.
However, we were not interested in bringing to bear the entire collection of machine learning algorithms to this prediction task in order to completely optimize a model.
Rather, we were interested in evaluating how well two cognitive models could explain the results, and then making selective modifications to them to see how performance changes.
The modifications that we explored were selected such that they influenced model accuracy, could be easily added to (or removed from) the cognitive models without massive modifications, and could be considered cognitively plausible.

\subsection{Research Questions}

The core question motivating this research is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site selects?
This question can be tested by generating a prediction for the most likely selected tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag selection \parencite{Efron2010} and StackOverflow tag selection \parencites{Stanley2013}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

\subsubsection{StackOverflow and Twitter}

These two memory models were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

The two memory retrieval models were also tested on tag selection for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and were interested in retrieval models that generalize across tasks.
However, they are also similar on several accounts:
Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags for Twitter and tags for StackOverflow),
and the user data from the datasets are publicly available for analysis.

\subsection{Roadmap}

This rest of this document is separated into 5 sections.
``Theoretical Background'' describes the vector-based and Bayesian declarative memory models that were evaluated in this research,
shows how these models are related to the other common declarative memory models,
and identifies examples of similar tagging models that have already been applied to various tag recommendation systems.
``General Methodology'' presents the overall method for testing the models on the StackOverflow and Twitter datasets.
``Past User Behavior'' shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
``Combining Predictors'' adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
The final section provides a general discussion of the overall findings.

\section{Theoretical Background}

\subsection{ACT-R Declarative Memory Theory}

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
At the core of the system is the concept of a ``chunk'': the representation of a piece of knowledge (e.g., a word).
The equations that make up this system were derived from a rational analysis of declarative memory retrieval \parencite{Anderson1989}.
That is, given the task of retrieving a chunk of information from declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

\subsubsection{ACT-R DM Model}

A formal description of the ACT-R Declarative Memory model is included in Table \ref{tabACTRModel}.

\begin{table}[!ht]
  \caption{ACT-R declarative memory model}
  \label{tabACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
      \hline
    \end{tabu}
  }
\end{table}

The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.

\subsubsection{Base-Level Activation}

Base-level activation reflects the log prior odds of needing an observed chunk again.
The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.

\subsubsection{Strength of Association}

Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
the likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.

\subsubsection{Connection to Pointwise Mutual Information}

Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
The PMI index is presented as Equation \eqref{eqPMI}.

\begin{equation}
  \label{eqPMI}
  \mathit{PMI}(y,x) = log \frac{p(y|x)}{p(y)}
\end{equation}

Further, this PMI equation is identical to ACT-R's strength of association ($S_{ji}$) equation for large datasets.
Once the number of occurrences is large enough such that the probability of observing any particular contextual element ($p(x)$) is near zero, then ACT-R's strength of association index simplifies to the PMI equation.

\subsubsection{Attentional Weight}

The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.

\subsection{Vector-Based Memory Systems}

``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems are an alternative representation of declarative memory.
They have been successfully used to predict memory retrievals that leverage word order information \parencite{Jones2007}.
They also have similar structure to memory models that use Latent Semantic Analysis \parencite{Landauer1997} (LSA).

Vector-based models represent a concept (i.e., a chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.

For vector-based systems, each word is represented by an environment vector $e_{i}$ that is nearly orthogonal to all other words' environment vectors.
The number of dimensions for these vectors is much less than the number of rows in a full word co-occurrence matrix (typically around 2,000 - 4,000), which is how vector-based systems represent information in a much more compact space.
Paired with each environment vector ($e_{i}$) is a memory vector ($m_{i}$) that contains the summed representation of all other environment vectors that have co-occurred with that $e_{i}$ environment vector.
These memory vectors can contain environment vectors from position-independent word co-occurrence information ($c_{i}$) and word order information ($o_{i}$), which can be combined into a single representation ($m_{i}$).
Over time, memory vectors accumulate a distributed representation of the most common environment vectors that co-occurred with them.

\subsubsection{Retrieval Process}

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.

Resonance is the opposite retrieval process, where a memory vector is created from the context, and then that context memory vector is compared to all memory vectors.
For example, assume that a user has written the word ``zend'' in a post, and she wants to assign a tag for the post.
``zend'' has co-occurred with the tag \emph{PHP} many times previously, so the unordered memory vector for ``PHP'' ($c_{\mathit{PHP}}$) contains the unordered environment vector for \emph{zend} ($e_{zend}$).
To determine the most correlated memory vector with context, a memory vector is created from the current context ($e_{zend}$).
The cosine between that context memory vector and the memory vector for \emph{PHP} will be high, since the \emph{PHP} memory vector ($c_{\mathit{PHP}}$) often contains the context environment vector ($e_{zend}$).
Since the cosine is high, the tag \emph{PHP} will most likely be returned as the vector that resonates highest with the current context.

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.

\subsubsection{Word Order for Vector-Based Systems}

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
With vector-based systems, order information is included by creating an environment vector ($e_{bind(i,\Phi)}$) that is a function of environment vectors for both the word ($e_{i}$) and location ($\Phi$).
This function has the useful property that the resulting environment vector and the location vector can be inverted to return the original word's environment vector.
Circular convolution and inverse circular convolution (i.e., circular correlation) are used by \textcite{Plate1995} and \textcite{Jones2007}'s BEAGLE model to implement these operations.
This allows for retrieval requests such as ``playing \#$\Phi$,'' ``just read \#$\Phi$,'' and ``vector-based $\Phi$ systems.''

\subsubsection{Random Permutation Model}

In order to represent word order information, a function must be used that converts an environment vector for a set of words and positions (e.g., ``stack $\Phi$'') into a new uncorrelated environment vector.
\textcite{Sahlgren2008} used a simple random permutation method for this operation.

A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.

\begin{table}[!ht]
  \caption{Random permutation model}
  \label{tabRandPermModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &		$A_{i} = r(m_{C},m_{i})$ \\
      Memory Vector &		$m_{i} = \sum_{i \in all past} c_{i} + \sum_{i \in all past} \sum_{l \in locations} o_{i,l}$ \\
      Unordered Context &	$c_{i} = e_{i}$ \\
      Ordered Context &		$o_{i,l} = e_{i^{-l}}$ \\
      Context Memory Vector &	$m_{C} = \sum_{i \in C} c_{i} + \sum_{i \in C} \sum_{l \in locations} o_{i,l}$ \\
      Environment Vector & 	$e_{i} = rand$ \\
      \hline
    \end{tabu}
  }
\end{table}

\subsubsection{Connection to LSA}

Both vector-based models and LSA calculate word similarity on a reduced co-occurrence matrix.
For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
LSA uses Singular Value Decomposition (SVD) to find words that have highly correlated document occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
For vector-based models, each tag (document) is represented by a compressed vector with a length much less than the number of words.
A vector-based approach like random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
Nonetheless, vector-based models have been shown to perform similar to if not better than LSA \parencites{Sahlgren2008,Jones2007}.

\subsection{Comparison of Vector-Based Models and ACT-R}

Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
\citeauthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect \parencite{Anderson1974}.

\subsubsection{Base-Level Activation}

Vector-based models and ACT-R's Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.

The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
The recent content in tweets on a news site like Twitter should be more relevant, for example.
\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.

ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.

\subsubsection{Word Order}

One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \noparencite{Lewis2006}).

\subsubsection{Activation Calculation}

Vector-based retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely correlated with one another.

>>>>

\documentclass[answers,12pt]{exam}
\usepackage{xcolor}
\definecolor{SolutionColor}{rgb}{0.1,0.3,1}
\usepackage{lipsum}
\usepackage{enumerate}
\usepackage{alphalph}

\renewcommand{\theenumi}{\AlphAlph{\value{enumi}}}
\renewcommand{\thequestion}{\AlphAlph{\value{question}}}
\renewcommand\questionlabel{\llap{\thequestion)}}

%\pointsinrightmargin
%\boxedpoints
\unframedsolutions
\shadedsolutions
\definecolor{SolutionColor}{rgb}{0.9,0.9,1}
\renewcommand{\solutiontitle}{}

@@

@@

\begin{questions}

\question

 The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
  By framing the task of choosing a hashtag for a tweet and tagging a post on StackOverflow as a declarative memory retrieval problem,
  two cognitively-plausible declarative memory models were applied to millions of posts and tweets and evaluated on how accurately they predict a user's chosen tags.
  An ACT-R based Bayesian model and a random permutation vector-based model were tested on the large datasets.
  The results show that past user behavior of tag use is a strong predictor of future behavior.
  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to an entropy-weighting natural language processing method used to attenuate high-frequency words (e.g., articles and prepositions).
  Word order was not found to be a strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.

\begin{solution}

The growth of social media and user-created content on online sites provides unique opportunities to study models of long-term memory.
By framing the task of choosing a hashtag for a tweet and tagging a post on StackOverflow as a long-term memory retrieval problem,
two long-term memory models were tested on millions of posts and tweets and evaluated on how accurately they predict a user's chosen tags.
An uncompressed and compressed model (in terms of storage of information in long-term memory) were tested on the large datasets.
The results show that past user behavior of tag use is a strong predictor of future behavior.
Furthermore, past behavior was successfully incorporated into the compressed model that previously used only context.
Also, an attentional weight term in the uncompressed model was linked to a natural language processing method used to attenuate common words (e.g., articles and prepositions).
Word order was not found to be a strong predictor of tag use, and the compressed model performed comparably to the uncompressed model without including word order.
This shows that the strength of the compressed model is not in the ability to represent word order, but rather in the way in which information is efficiently compressed.
The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.

\end{solution}

\end{questions}
